---
title: "Fyw estimation using traditional procedure"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fyw estimation using traditional procedure}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
# Load require packages
library(Fyw)
library(lubridate)
#library(ggplot2)  # uncomment to see plot
```

### 1. Overview
This document describes how the young water fraction is calculated based on the 
traditional procedure (Kirchner, 2016). 

First we need data for this demonstration. Let's take the isotope data from the Alp catchment.

```{r, message=FALSE}
# Get isotope in precipitation (P) and streamflow (S) from the the Alp catchment
isotopeP <- subset(isotopeData, catchment == "Alp" & variable == "precipitation")
isotopeS <- subset(isotopeData, catchment == "Alp" &  variable == "streamflow")
```

### 2. Fyw estimation using the traditional procedure
In the traditional procedure (Kirchner, 2016), Fyw is estimated using the following steps:

-   Step 1: Fit the sine wave to observed isotope in precipitation and streamflow 
using the Iteratively Least Squares (IRLS) regression approach. The fitted sine 
wave has the following form (please see Eq. 5; Kirchner, (2016)):

$$ c_P = a_P \cdot cos(2 \pi t) + b_P \cdot sin(2 \pi t) + k_P $$

```{r, message=FALSE}
#------------------------------------------------------------------------------#
#             Fit sine wave to observed isotope in precipitation               #
#------------------------------------------------------------------------------#
# First, convert date to decimal, ranging from [0, 1]
t <- decimal_date(isotopeP$date) - trunc(decimal_date(isotopeP$date))
  
# Fit to sine wave to observed isotope in P using IRLS method
sineP <- IRLS(Y = isotopeP$delta_18O,
              X = data.frame(cos = cos(2*pi*t), sin = sin(2*pi*t)),
              pweights = isotopeP$water_flux_mm)

# Uncomment to see the plot
# ggplot()+
#   geom_point(data = isotopeP, aes(x = date, y = delta_18O, color = "cP (observed)"), alpha = 0.5) + 
#   geom_line(aes(x = isotopeP$date, y = as.numeric(sineP$fitted.values), color = "cP (fitted sine-wave)"), linewidth = 0.75)

#------------------------------------------------------------------------------#
#             Fit sine wave to observed isotope in streamflow                  #
#------------------------------------------------------------------------------#
# First, convert date to decimal, ranging from [0, 1]
t <- decimal_date(isotopeS$date) - trunc(decimal_date(isotopeS$date))
  
# Fit to sine wave to observed isotope in P using IRLS method
sineS <- IRLS(Y = isotopeS$delta_18O,
              X = data.frame(cos = cos(2*pi*t), sin = sin(2*pi*t)),
              pweights = isotopeS$water_flux_mm)

# Uncomment to see the plot
# ggplot()+
#   geom_point(data = isotopeS, aes(x = date, y = delta_18O, color = "cS (observed)"), alpha = 0.5) + 
#   geom_line(aes(x = isotopeS$date, y = as.numeric(sineS$fitted.values), color = "cS (fitted sine-wave)"), linewidth = 0.75)
```

-   Step 2: Find Fyw from the two fitted sine waves
```{r, message=FALSE}
# Parameters of the fitted sine wave to isotope in precipitation
kP <- as.numeric(sineP$coefficients[1])
aP <- as.numeric(sineP$coefficients[2])
bP <- as.numeric(sineP$coefficients[3])

# Parameters of the fitted sine wave to isotope in precipitation
kS <- as.numeric(sineS$coefficients[1])
aS <- as.numeric(sineS$coefficients[2])
bS <- as.numeric(sineS$coefficients[3])

# Amplitudes of the fitted sine wave to precipitation and streamflow (please see Eq. 6; Kirchner, (2016))
AP <- sqrt(aP^2 + bP^2)
AS <- sqrt(aS^2 + bS^2)

# Phase and phase shift
phi <- findPhiPhaseShift(aP = aP, bP = bP, aS = aS, bS = bS)
phiP <- phi$phiP
phiS <- phi$phiS
phiS_phiP <- phi$phiS_phiP

# Parameters of the gamma (transit time distribution) (please see Eqs. 10, 11; Kirchner, (2016))
alphaBetaMTT <- findAlphaBetaMTT(phiS_phiP = phiS_phiP,
                                 AS = AS,
                                 AP = AP)

# Age threshold of the young water fraction (please see Eq. 14; Kirchner, (2016)))
tauyw <- findtauyw(alpha = alphaBetaMTT$alpha, method = "approximation")$tauyw

# Young water fraction (equals to the amplitude ratio)
Fyw_1 <- AS/AP

# Or the young water fraction can be calculated using the lower incomplete gamma function (please see Eq. 13; Kirchner, (2016))
Fyw_2 <- pgamma(q = tauyw,
                shape = alphaBetaMTT$alpha,
                scale = alphaBetaMTT$beta,
                lower.tail = TRUE)

# Show the results in data frame
t(data.frame(AP = AP, phiP = phiP, kP = kP, AS = AS, phiS = phiS, kS = kS, 
             Fyw_1 = Fyw_1, Fyw_2 = Fyw_2, tauyw = tauyw))
```
### References

Kirchner, J. W. 2016. “Aggregation in Environmental Systems – Part 1: Seasonal 
Tracer Cycles Quantify Young Water Fractions, but Not Mean Transit Times, in 
Spatially Heterogeneous Catchments.” Hydrology and Earth System Sciences 20 (1): 279–97. https://doi.org/10.5194/hess-20-279-2016.




