---
title: "Fyw estimation using the revised procedure"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fyw_revised_procedure}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, message=FALSE}
# Load require packages
library(Fyw)
library(ggplot2)
# devtools::install(build_vignettes = TRUE)
```

### 1. Overview
This document describes how the young water fraction is calculated based on the 
revised procedure (from this package). 

First we need data for this demonstration. Let's take the isotope data from the Alp catchment.

```{r, message=FALSE}
# Get isotope in precipitation (P) and streamflow (S) from the the Alp catchment
isotopeP <- subset(isotopeData, catchment == "Alp" & variable == "precipitation")
isotopeS <- subset(isotopeData, catchment == "Alp" &  variable == "streamflow")
```

### 2. Fyw estimation using the revised procedure
In the revised procedure, Fyw is estimated using the following steps:

-   **Step 1**: Fit the sine wave to observed isotope in precipitation. The fitted sine 
wave has the following form:


$$ c_P = A_P \cdot sin(2 \pi t + \varphi_P)  + k_P $$

```{r, message=FALSE, fig.height=2.5, fig.width=7}
#------------------------------------------------------------------------------#
#             Fit sine wave to observed isotope in precipitation               #
#------------------------------------------------------------------------------#
fitSineP <- fitSineNL(observed = isotopeP$delta_18O,
                      a = c(1,3),
                      phi = c(0, 2*pi),
                      k = c(-12,-8), 
                      t = isotopeP$date,
                      nIter = 10000,
                      nBestIter = 10,
                      weight = isotopeP$water_flux_mm,
                      setSeed = 2023)

# Plot the results (fitted sine wave and observed)
ggplot(fitSineP$simulated)+
  geom_line(aes(x = date, y = simulated, color = simulation)) +         # 10 best fitted sine waves
  geom_point(data = fitSineP$observed, aes(x = date, y = observed)) +   # observed data
  labs(x = "", y = expression(paste(delta^{18}, "O in precipitation (‰)"))) +
  theme(legend.position = "none")
```

-   Step 2: Find parameters of the gamma distribution (alpha, beta) by convoluting
the fitted sine wave to isotope in precipitation (step 1) with the gamma functions
and evalute the goodness-of-fit with observed isotope in streamflow. The following code
was not run because time-consuming.

```{r, message=FALSE, message=FALSE, fig.height=2.5, fig.width=7, eval = FALSE}
# NOTE: Due to high computational demand, here I set nIter = 4 and nWarmupYears = 5
#       but to have a better fit nIter should be large (e.g, 2000), 
#       and nWarmupYears should be large (e.g., 10) to reduce the impact of initial conditions
#       and the alphaRange, betaRange should be larger (e.g., from 0.01 to 5)

# Here we randomly combined 10 best fit sine wave functions (Step 1) with 3 randomly
#       generated gamma parameters and extract results of the 2 best fitted gamma functions

# Parameters of the fitted sine wave to isotope in precipitation
fitGamma <- fitGamma(AP = fitSineP$parameter$a,
                     phiP = fitSineP$parameter$phi,
                     kP = fitSineP$parameter$k,
                     alphaRange = c(0.75,0.9),
                     betaRange = c(0.2,0.3),
                     simulatedDate = isotopeS$date,
                     fittedData = isotopeS$delta_18O,
                     weight = isotopeS$water_flux_mm,
                     nIter = 3,
                     nBestIter = 2,
                     nCores = 1,
                     nWarmupYears = 5,
                     setSeed = 2023)

# Plot to see the quality of fit
ggplot()+
  geom_point(data = isotopeS, aes(x = date, y = delta_18O), alpha = 0.35) +                  # Observed
  geom_line(data = fitGamma$simulated, aes(x = date, y = simulated, color = simulation)) +   # Simulated
  labs(x = "", y = expression(paste(delta^{18}, "O in streamflow (‰)"))) +
  theme(legend.position = "none")
```

-   Step 3: Now calculate the young water fraction with user defined age threshold

```{r, message=FALSE, message=FALSE, fig.height=2.5, fig.width=7, eval = FALSE}

# Let's say we want to find the Fyw fraction with age threshold of 2 months
tauyw <- 2/12 

# Find Fyw 
Fywater <- c()
for (i in 1:nrow(fitGamma$parameterSet)){
  Fywater <- c(Fywater, pgamma(q = tauyw,
                               shape = fitGamma$parameterSet$alpha[i],
                               scale = fitGamma$parameterSet$beta[i],
                               lower.tail = TRUE))
}

# show Fyw values
Fywater

```

